---
output:
  md_document:
    variant: gfm 
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# gmat
An R package for simulating correlation matrices possibly constrained by acyclic
directed and undirected graphs.

[![Build Status](https://travis-ci.com/irenecrsn/gmat.svg?branch=master)](https://travis-ci.com/irenecrsn/gmat)
[![codecov](https://codecov.io/gh/irenecrsn/gmat/branch/dev/graph/badge.svg)](https://codecov.io/gh/irenecrsn/gmat)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/gmat)](https://CRAN.R-project.org/package=gmat)
[![CRAN status](http://www.r-pkg.org/badges/version/gmat)](https://CRAN.R-project.org/package=gmat)

## Installation

The package is available on `CRAN`, to get the latest stable version use:
```{r, eval = FALSE}
install.packages("gmat")
```

Alternatively, using the R package `devtools` one may install the development version:
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("irenecrsn/gmat")
```

The other R packages required for `gmat` are `igraph` and `gRbase`, which can
also be installed from CRAN.

## Overview

The package mostly implements methods described in the following papers:

* Córdoba I., Varando G., Bielza C., Larrañaga P. __A fast Metropolis-Hastings
  method for generating random correlation matrices__. _Lecture Notes in Computer
  Science_ (IDEAL 2018), vol 11314, pp. 117-124, 2018.
* Córdoba I., Varando G., Bielza C., Larrañaga P. __A partial orthogonalization
  method for simulating covariance and concentration graph matrices__.
  _Proceedings of Machine Learning Research_ (PGM 2018), vol 72, pp. 61-72, 2018.
* Córdoba I., Varando G., Bielza C., Larrañaga P. __Generating random Gaussian
  graphical models__. _arXiv:1909.01062_, 2019.

## An example of use 

First, we generate a random undirected graph with `3` nodes and density `0.5`.
Then we generate, using our `port()` function, `2` correlation matrices
consistent with such random graphical structure.

```{r}
ug <- gmat::rgraph(p = 3, d = 0.5)
igraph::print.igraph(ug)
gmat::port(N = 2, ug = ug)
```

We appreciate how the zero pattern is shared by all of the simulated matrices.
The return value is an array, and so the individual matrices can be accessed as
`matrices[, , n]`, where `n` is the index of the matrix we want to retrieve from
the sample, ranging from `1` to `N`.

We may also sample correlation matrices using i.i.d.  coefficients in their
upper Cholesky factor `U`.
```{r}
gmat::chol_iid(N = 2)
```
A specific zero pattern can be enforced in `U` using an acyclic digraph.
```{r}
dag <- gmat::rgraph(p = 3, d = 0.5, dag = TRUE)
igraph::print.igraph(dag)
top_sort <- igraph::topo_sort(dag)
peo_sort <- rev(top_sort)
inv_top <- order(top_sort)
inv_peo <- order(peo_sort)
m <- gmat::chol_iid(dag = dag)[, , 1]
# We sort the matrix according to both a topological and a perfect elimination
# ordering
m_top_sort <- m[top_sort, top_sort]
m_peo_sort <- m[peo_sort, peo_sort]
# Transpose of Cholesky factor (lower Cholesky decomposition)
U_chol <- chol(m_peo_sort) 
# Upper Cholesky factor (upper Cholesky decomposition)
U <- gmat::uchol(m_top_sort)
print(U_chol[inv_peo, inv_peo])
print(U[inv_top, inv_top])
print(m)
```
The zeros are correctly reflected in the upper Cholesky factor when using a
topological sorting, whereas they are reflected in the standard lower Cholesky
decomposition factor if we sort the matrix with a perfect elimination ordering.

See more examples and paper references at [the documentation
website](https://irenecrsn.github.io/gmat/) for the package.
